---
title: "FIFA Soccer Best Team Formation"
author: "Mohit Abbi"
date: "5/15/2019"
output: html_document
---

# Installing and intializing the necessary libraries
install.packages("car")

```{r}
library(MASS)
library(car)
```

#Loading CompleteDataset

```{r}
data1 <- read.csv("C:/Users/Dell-pc/Desktop/Github/Projects/FIFA Soccer Best Team Formation/CompleteDataset.csv")
head(data1)
str(data1)
View(data1)
colnames(data1)
```

# Data Cleaning

#Coverting factor variables

```{r}
factor(data1$Nationality)
a <- table(data1$Nationality)

triala = as.data.frame(a)
#View(triala)
sort(triala$Freq, decreasing=TRUE)
triala$freq_percent = (triala$Freq / sum(triala$Freq))*100
triala$dummy  <- as.numeric(triala$freq_percent >=5)


for (i in 1:nrow(triala)){ 
  if(triala$dummy == 0) triala$dummy2[i]=c("other") 
  else 
    triala$dummy2[i] == triala$Var1[i]
} 

#Removing all the unnecessary variables from the dataset
data2 <- subset(data1, select = -c(Name, Photo, Flag, Club.Logo, Value, Wage, Preferred.Positions,Nationality,Club))
str(data2)
#View(data2)

# Convert factor variables to integer
typeconversion <- function(columnname) {as.integer(levels(columnname))[columnname]}
df1 <- lapply(data2[c(6:39)], FUN = typeconversion)
#View(df1)
str(df1)

#Convert list to dataframe
df2 <- as.data.frame(df1)
df2 <- cbind(data2[c(1:5,40:66)], df2)
#View(df2)
str(df2)

#Here, we filled all the blanks with NA and omitted all the NA's from the data
df2[df2 == ""] <- NA
df2 <- na.omit(df2)
```

#Divide the data into train and test data

```{r}
set.seed(1) 
row.number <- sample(x=1:nrow(df2), size=0.7*nrow(df2))
train = df2[row.number,]
test = df2[-row.number,]
head(train)
head(test)
```

##Optimization

#PCA 
#prepare data for pca

```{r}
pca.data <- subset(df2, select = -c(3))
df2.pca <- prcomp(pca.data, retx = TRUE, center = TRUE, scale. = TRUE)
summary(df2.pca)
#df2.pca$x

pcamodeldata <- as.data.frame(cbind(df2$Overall,(df2.pca$x)))
row.number <- sample(x=1:nrow(pcamodeldata), size=0.7*nrow(pcamodeldata))
pcatrain = pcamodeldata[row.number,]
pcatest = pcamodeldata[-row.number,]
head(pcatrain)
head(pcatest)
```

##Linear Regression on PCA Components

```{r}
pcamodel <- lm(formula = V1~., data = pcatrain)
summary(pcamodel)

##Model Validation - VIF
vif(pcamodel)

##Linear pca model with "stepAIC"
steppca <- step(pcamodel, direction="backward")
steppca$anova # display results
summary(steppca)
vif(steppca)

pcatestmodel <- predict(steppca, newdata = pcatest)
summary(pcatestmodel)

observedval <- pcatest$V1
```

# Observing Errors 

```{r}
SSE <- sum((observedval - pcatestmodel) ^ 2)
SST <- sum((observedval - mean(observedval)) ^ 2)
r2 <- 1 - SSE/SST
r2

#RMSE:
rmse <- sqrt(sum((pcatestmodel - observedval)^2)/length(observedval))
rmse

#MSE
mse <- (sum((pcatestmodel - observedval)^2)/length(observedval))
mse

max(pcatestmodel - observedval)
min(pcatestmodel - observedval)
```

##Base Model1
##Linear Regression on the whole dataset
 #View(train)

```{r}
model1 <- lm(formula = Overall ~ ., data = train)
summary(model1)

##Model Validation - VIF
alias(lm(formula = Overall ~ ., data = train))
# vif(model1)

#Fit the model to test dataset
testmodel <- predict(model1, newdata = test)
summary(testmodel)

observedval <- test$Overall
```

#Compute the error values

```{r}
SSE <- sum((observedval - testmodel) ^ 2)
SST <- sum((observedval - mean(observedval)) ^ 2)
r2 <- 1 - SSE/SST
r2

#RMSE:
rmse <- sqrt(sum((testmodel - observedval)^2)/length(observedval))
rmse

#MSE
mse <- (sum((testmodel - observedval)^2)/length(observedval))
mse
```

##Model Comparison (Selection Method - AIC)
#Applying AIC method on model1

```{r}
step1 <- stepAIC(model1, direction = "both")
step1$anova 

model1aic <- lm(formula = Overall ~ X + Age + Potential + CM + ID + LB + LM + LS + Acceleration + 
                  Aggression + Balance + Ball.control + Composure + Crossing + 
                  Finishing + GK.positioning + Long.shots + Positioning + Reactions + 
                  Short.passing + Shot.power + Sliding.tackle + Stamina + Standing.tackle + 
                  Strength, data = train)
summary(model1aic)

##Model Validation - VIF
vif(model1aic)

#Fit the model to test aic dataset
testmodel1aic <- predict(model1aic, newdata = test)
summary(testmodel1aic)

observedval <- test$Overall
```

#Compute the error values
```{r}
SSE <- sum((observedval - testmodel1aic) ^ 2)
SST <- sum((observedval - mean(observedval)) ^ 2)
r2 <- 1 - SSE/SST
r2

#RMSE:
rmse <- sqrt(sum((testmodel1aic - observedval)^2)/length(observedval))
rmse

#MSE
mse <- (sum((testmodel1aic - observedval)^2)/length(observedval))
mse

max(testmodel1aic - observedval)
min(testmodel1aic - observedval)
```

##Base Model2
##Linear model after removing insignificant variables

```{r}
model2 <- lm(formula = Overall ~ X + Age + Potential + CM + LB + LS + Acceleration
             + Balance + Ball.control + Composure + Reactions + Short.passing + Sliding.tackle+
               Stamina + Strength, data = train)
summary(model2)

testmodel2 <- predict(model2, newdata = test)
summary(testmodel2)

##Model Validation - VIF
vif(model2)

observedval <- test$Overall
```
#Compute the error values

```{r}
SSE <- sum((observedval - testmodel2) ^ 2)
SST <- sum((observedval - mean(observedval)) ^ 2)
r2 <- 1 - SSE/SST
r2

#RMSE:
rmse <- sqrt(sum((testmodel2 - observedval)^2)/length(observedval))
rmse

#MSE
mse <- (sum((testmodel2 - observedval)^2)/length(observedval))
mse

max(testmodel2 - observedval)
min(testmodel2 - observedval)
```

##Model Comparison (Selection Method - AIC)
#Applying AIC method on model2

```{r}
step2 <- stepAIC(model2, direction = "both")
step2$anova 

model2aic <- lm(formula = Overall ~ X + Age + Potential + CM + LB + LS + Acceleration + 
                  Balance + Ball.control + Composure + Reactions + Short.passing + 
                  Sliding.tackle + Stamina + Strength, data = train)
summary(model2aic)

#Fit the model to test aic dataset
testmodel2aic <- predict(model2aic, newdata = test)
summary(testmodel2aic)

observedval <- test$Overall
```

#Compute the error values

```{r}
SSE <- sum((observedval - testmodel2aic) ^ 2)
SST <- sum((observedval - mean(observedval)) ^ 2)
r2 <- 1 - SSE/SST
r2

#RMSE:
rmse <- sqrt(sum((testmodel2aic - observedval)^2)/length(observedval))
rmse

#MSE
mse <- (sum((testmodel2aic - observedval)^2)/length(observedval))
mse

max(testmodel2aic - observedval)
min(testmodel2aic - observedval)
```

##Base Model3
##Linear model after including only skills and not position data

```{r}
 #View(train)
modeldata1 <- subset(train, select = -c(1:2,4:33))
 #View(modeldata1)
model3 <- lm(formula = Overall ~ ., data = modeldata1)
summary(model3)

testmodel3 <- predict(model3, newdata = test)
summary(testmodel3)

##Model Validation - VIF
vif(model3)
observedval <- test$Overall
```
#Compute the error values

```{r}
SSE <- sum((observedval - testmodel3) ^ 2)
SST <- sum((observedval - mean(observedval)) ^ 2)
r2 <- 1 - SSE/SST
r2

#RMSE:
rmse <- sqrt(sum((testmodel3 - observedval)^2)/length(observedval))
rmse

#MSE
mse <- (sum((testmodel3 - observedval)^2)/length(observedval))
mse

max(testmodel3 - observedval)
min(testmodel3 - observedval)
```
##Model Comparison (Selection Method - AIC)  
#Applying AIC Method on model3

```{r}
step3 <- stepAIC(model3, direction = "both")
step3$anova 

model3aic <- lm(formula = Overall ~ Balance + Ball.control + Composure + Crossing + Curve + 
                  Dribbling + Finishing + Free.kick.accuracy + Heading.accuracy + 
                  Interceptions + Jumping + Long.passing + Long.shots + Marking + 
                  Penalties + Positioning + Reactions + Short.passing + Shot.power + 
                  Sliding.tackle + Sprint.speed + Stamina + Standing.tackle + 
                  Strength + Vision, data = train)
summary(model3aic)

#Fit the model to test aic dataset
testmodel3aic <- predict(model3aic, newdata = test)
summary(testmodel3aic)

##Model Validation - VIF
vif(model3aic)

observedval <- test$Overall
```

#Compute the error values
```{r}
SSE <- sum((observedval - testmodel3aic) ^ 2)
SST <- sum((observedval - mean(observedval)) ^ 2)
r2 <- 1 - SSE/SST
r2

#RMSE:
rmse <- sqrt(sum((testmodel3aic - observedval)^2)/length(observedval))
rmse

#MSE
mse <- (sum((testmodel3aic - observedval)^2)/length(observedval))
mse

max(testmodel3aic - observedval)
min(testmodel3aic - observedval)
```
##Base Model4
##Linear model after including only skills and not position data

```{r}
 #View(train)

modeldata1 <- subset(train, select = -c(1:2,4:33))
 #View(modeldata1)
model4 <- lm(formula = Overall ~ Balance + Ball.control + Composure + Crossing + Curve + Finishing +
               + Free.kick.accuracy + Heading.accuracy + Interceptions + Jumping + Long.passing
             + Long.shots + Penalties + Positioning + Reactions + Short.passing + Sliding.tackle
             + Shot.power + Stamina + Standing.tackle + Strength + Stamina + Sprint.speed + Vision, data = modeldata1)
summary(model4)

testmodel4 <- predict(model4, newdata = test)
summary(testmodel4)

##Model Validation - VIF
vif(model4)

observedval <- test$Overall
```
#Compute the error values
```{r}
SSE4 <- sum((observedval - testmodel4) ^ 2)
SST4 <- sum((observedval - mean(observedval)) ^ 2)
r24 <- 1 - SSE4/SST4
r24

#RMSE:
rmse4 <- sqrt(sum((testmodel4 - observedval)^2)/length(observedval))
rmse4

#MSE
mse4 <- (sum((testmodel4 - observedval)^2)/length(observedval))
mse4

max(testmodel4 - observedval)
min(testmodel4 - observedval)
```
##Model Comparison (Selection Method - AIC)
#Applying AIC Method on model4

```{r}
step4 <- stepAIC(model4, direction = "both")
step4$anova 

model4aic <- lm(formula = Overall ~ Balance + Ball.control + Composure + Crossing + Curve + 
                  Finishing + +Free.kick.accuracy + Heading.accuracy + Interceptions + 
                  Jumping + Long.passing + Long.shots + Penalties + Positioning + 
                  Reactions + Short.passing + Sliding.tackle + Shot.power + 
                  Stamina + Standing.tackle + Strength + Stamina + Sprint.speed + 
                  Vision, data = train)
summary(model4aic)

#Fit the model to test aic dataset
testmodel4aic <- predict(model4aic, newdata = test)
summary(testmodel4aic)

##Model Validation - VIF
vif(model4aic)

observedval <- test$Overall
```
#Compute the error values

```{r}
SSE <- sum((observedval - testmodel4aic) ^ 2)
SST <- sum((observedval - mean(observedval)) ^ 2)
r2 <- 1 - SSE/SST
r2

#RMSE:
rmse <- sqrt(sum((testmodel4aic - observedval)^2)/length(observedval))
rmse

#MSE
mse <- (sum((testmodel4aic - observedval)^2)/length(observedval))
mse

max(testmodel4aic - observedval)
min(testmodel4aic - observedval)
```
